---
title: "Example from class"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

# install.packages("devtools")
devtools::install_github("thomasp85/patchwork")

library(ggplot2)
library(patchwork)

p3 <- ggplot(mtcars) + geom_smooth(aes(disp, qsec))
p4 <- ggplot(mtcars) + geom_bar(aes(carb))

(p1 | p2 | p3) /
      p4

```

```{r}

library(tidyverse)
 library(ggplot2)
data("mtcars")

library(patchwork)

devtools::install_github("thomasp85/patchwork")

p3 <- ggplot(mtcars) + geom_smooth(aes(disp, qsec))
p4 <- ggplot(mtcars) + geom_bar(aes(carb))

(p1 | p2 | p3) /
      p4

(p1 | p2 | p3)

 
mtcars$gear <- as.factor(mtcars$gear)
model <- lm(mpg ~ wt * gear, data = mtcars)

model

```

*(Intercept)*: 28.395
Our intercept for a car with a weight of 0 and 3 gears (because that is our first level)

*wt*: -3.157 
when we increase weight by one unit, we decrease mpg by -3.1157 for g3.

*gear4*: main effect, what we have to add to intercept (combine this with intercept will give us a car with 4 gears and weight zero)
If you want the slope for group 4, you add wt:gear4 and wt. -> 

*gear5*: 14.168 

*wt:gear4*: -3.707 
interaction!
how much u should adjust the slope. how it differs (add 3.707 with 3.157)

*wt:gear5*: -4.889 


If you want the slope for group 4, you add wt:Gear and wt.
If you want intercept for gear4, you add intercept with gear4.


# Below 

```{r}

mtcars$int <- interaction(mtcars$gear, mtcars$wt) # estimate. you create a new variable of all possible combinations of the levels.

```
 

# Interaction plot

```{r}

class(mtcars$gear)
interaction.plot(model)
```



